---
title: "Median co-expression of PD1 VS PD-L1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Using mRNA expression data from TCGA identify the
disease type with the highest median co-expression of PD-1 and PD-L1.

* PDCD1 = gene for PD-1

* CD274 = gene for PD-L1

```{r cohorts, echo = FALSE}
library(FirebrowseR)
diff.Exp.Genes = c("PDCD1","CD274")
all.Found = F
page.Counter = 1
mRNA.Exp = list()
page.Size = 2000 # using a bigger page size is faster
#https://gdc.cancer.gov/resources-tcga-users/tcga-code-tables/sample-type-codes
while(all.Found == F){
  # select primary solid tumors
  mRNA.Exp[[page.Counter]] = Samples.mRNASeq(format = "csv", protocol="RSEM", sample_type=c("TP"),
                                             gene = diff.Exp.Genes,
                                             page_size = page.Size,
                                             page = page.Counter)
  if(page.Counter > 1)
    colnames(mRNA.Exp[[page.Counter]]) = colnames(mRNA.Exp[[page.Counter-1]])
  
  if(nrow(mRNA.Exp[[page.Counter]]) < page.Size)
    all.Found = T
  else
    page.Counter = page.Counter + 1
}
mRNA.Exp.tot = do.call(rbind, mRNA.Exp)
#dim(mRNA.Exp.tot)
mRNA.Exp.complete = subset( mRNA.Exp.tot, expression_log2!='None')
#dim(mRNA.Exp.complete)

```

```{r plot, echo = FALSE}

pd1 = subset(mRNA.Exp.complete, gene=="PDCD1")
pd1$expression_log2 = as.numeric(pd1$expression_log2)
median_pd1 = with(pd1, tapply(expression_log2, cohort, median))
#median_pd1

pdl1 = subset(mRNA.Exp.complete, gene=="CD274")
pdl1$expression_log2 = as.numeric(pdl1$expression_log2)
median_pdl1 = with(pdl1, tapply(expression_log2, cohort, median))
#median_pdl1

table.Exp = data.frame( cohort=names(median_pd1), pd1=median_pd1, pdl1=median_pdl1)

#names(median_pdl1) == names(median_pd1)

library(ggplot2)
plot( median_pd1 ~ median_pdl1, data=table.Exp, main="Median Expression PD1 vs PD-L1 per Cohort", xlab="Median Expression PD-L1/ Cohort", ylab="Median Expression PD1/ Cohort", col="red", pch=20, cex=2 )
text( median_pd1 ~ median_pdl1, data=table.Exp, labels=rownames(table.Exp), cex=0.7, font=2)
```

```{r table, echo = FALSE}
library(DT)
table.Exp.sorted = table.Exp[ with(table.Exp, order( c(median_pd1+median_pdl1),decreasing=T)),]
write.table(table.Exp.sorted, file = "PD1-vs-PD-L1.tsv", sep = "\t")

datatable(table.Exp.sorted, rownames=F, colnames=c("Cohort","Median Expression of PD1/Cohort","Median Expression of PD-L1/Cohort"),
          caption="Median Normalized RSEM expression for PD1 and PD-L1 per cohort sorted in decending order by sum of expression values.")

```

## Conclusion

The THYM cohort, solid tumor in the thymus, exhibits the highest co-expression of PD1 and PD-L1, followed by the DLBC (Diffuse Large B-cell Lymphoma). 

More information about these cohorts can be found at [firebrowse.org](http://firebrowse.org/).





